---
title: "BST210_project_priQ"
format: html
editor: visual
---

### Library
```{r}
library(tidyverse)   
library(haven)      
library(NHANES)     
library(dplyr)  
library(purrr)
library(lubridate)
library(glmnet)
library(vip)
library(car)
```

### Form Dataset
```{r}
getwd()
setwd("./data/2020")

ALQ <- read_xpt("P_ALQ.XPT")
BIOPRO <- read_xpt("P_BIOPRO.XPT")
BMX <- read_xpt("P_BMX.XPT")
BPXO <- read_xpt("P_BPXO.XPT")
DEMO <- read_xpt("P_DEMO.XPT")
#FOLATE <- read_xpt("P_FOLATE.XPT")
PAQ <- read_xpt("P_PAQ.XPT")
SLQ <- read_xpt("P_SLQ.XPT")
TCHOL <- read_xpt("P_TCHOL.XPT")
DPQ_raw <- read_xpt("P_DPQ.XPT")
FETIB <- read_xpt("P_FETIB.XPT")

DPQ <- DPQ_raw %>% 
    filter(complete.cases(select(., starts_with("DPQ"))[-1])) %>% 
    mutate(depression_score = rowSums(select(., starts_with("DPQ"))[-1])) |> #%>% 
    mutate(depression_category = case_when(
        depression_score <= 4 ~ "Minimal",
        depression_score <= 9 ~ "Mild",
        depression_score <= 14 ~ "Moderate",
        depression_score <= 19 ~ "Moderately severe",
        depression_score >= 20 ~ "Severe"
    ))

BPXO <- BPXO %>%
  mutate(
    MAP1 = (1 / 3) * BPXOSY1 + (2 / 3) * BPXODI1,
    MAP2 = (1 / 3) * BPXOSY2 + (2 / 3) * BPXODI2,
    MAP3 = (1 / 3) * BPXOSY3 + (2 / 3) * BPXODI3,
    avg_MAP = rowMeans(cbind(MAP1, MAP2, MAP3), na.rm = TRUE)
  )
datasets <- list(ALQ, BIOPRO, BMX, BPXO, DEMO,  PAQ, SLQ, TCHOL, DPQ)
test <- reduce(datasets, inner_join, by = "SEQN")
temp <- select(test, SEQN , BMXBMI, avg_MAP, ALQ130, RIAGENDR, RIDRETH1, RIDAGEYR, INDFMPIR, PAD680, LBDTCSI, depression_score, depression_category)

dat_raw <- left_join(SLQ, temp)

new_colnames <-  c(
  "ID",                          # SEQN; Respondent sequence number
  "sleep_time_weekdays",       # SLQ300;Usual_sleep_time_weekdays
  "wake_time_weekdays",        # SLQ310;Usual_wake_time_weekdays
  "sleep_hours_weekdays",            # SLD012;Sleep_hours_weekdays
  "sleep_time_weekends",       # SLQ320;Usual_sleep_time_weekends
  "wake_time_weekends",        # SLQ330;Usual_wake_time_weekends
  "sleep_hours_weekends",            # SLD013;Sleep_hours_weekends
  "frq_snore",                 # SLQ030;How_often_snore
  "frq_snort_or_stop_breathing",# SLQ040;How_often_snort_or_stop_breathing
  "sleep_trouble",  # SLQ050;Ever_told_doctor_sleep_trouble
  "overly_sleepy",          # SLQ120;Feel_overly_sleepy_day
  "BMI",                         # BMXBMI; Body Mass Index
  "avg_MAP",                     # avg_MAP; Average Mean Arterial Pressure
  "alcohol",                     # ALQ130; How often drink alcohol
  "gender",                      # RIAGENDR; Gender
  "Race_Hispanic_origin",        # RIDRETH1; Race/Hispanic origin
  "age",                         # RIDAGEYR; Age at screening
  "Family_income_to_poverty_ratio",  # INDFMPIR; Family income to poverty ratio
  "sedentary_activity",          # PAD680; Minutes sedentary activity
  "total_cholesterol",           # LBDTCSI; Total cholesterol mmol/L
  "depression_score",            # depression_score; Depression score
  "depression_category"          # depression_category; Depression category
)
colnames(dat_raw) <- new_colnames
#remove all lines with NA
dat_raw[dat_raw == ""] <- NA
dat_clean <- dat_raw |>
  drop_na() |>
   filter(alcohol < 16 & frq_snore != 7 & frq_snore != 9 & frq_snort_or_stop_breathing != 7 & frq_snort_or_stop_breathing != 9 & sleep_trouble != 9 & overly_sleepy != 9) |>
  mutate(across(c(sleep_time_weekdays,wake_time_weekdays,sleep_time_weekends,wake_time_weekends), hm)) |>
  mutate(across(c(ID,frq_snore,frq_snort_or_stop_breathing,sleep_trouble,overly_sleepy,gender,Race_Hispanic_origin),as.factor)) |>
  mutate(across(c(,),as.numeric)) |>
  mutate(sleep_hours_avg = 2/7*sleep_hours_weekends+5/7*sleep_hours_weekdays)
```

```{r}
pri_Q <- dat_clean[,c("sleep_trouble","BMI","avg_MAP","total_cholesterol","alcohol","gender","Race_Hispanic_origin","age","Family_income_to_poverty_ratio","depression_score","depression_category")]
pri_Q$sleep_trouble <- ifelse(dat_clean$sleep_trouble == "2",0,1)
```

```{r}
cate <- glm(sleep_trouble ~ as.factor(depression_category), data = pri_Q, family = "binomial")
conti <- glm(sleep_trouble ~ depression_score, data = pri_Q, family = "binomial")
summary(cate)
summary(conti)
anova(cate, conti, test="Chisq")
pri_Q <- pri_Q |>
  select(-depression_category)
```

```{r}
full_model <- glm(sleep_trouble ~ BMI+alcohol+gender+Race_Hispanic_origin+age+Family_income_to_poverty_ratio+depression_score+avg_MAP+total_cholesterol,data = pri_Q, family = "binomial")
vif(full_model)
```
```{r}
plot(conti$fitted.values~pri_Q$depression_score,type='p',
     col='black',ylab="P(sleep~disorder)",xlab="age")
lines(cate$fitted.values~pri_Q$depression_score,type='p',
      col="red")
legend('bottomright',legend=c("linear","ordinal"),
       pch=c(1,1),col=c(1,2))
```

```{r,eval=FALSE}
'''
set.seed(17)
lambda_grid <- .2 ^ (-50:50)
x <- model.matrix(sleep_trouble~., data=pri_Q)[,-c(1)]
y <- pri_Q$sleep_trouble
for (i in seq(0,1,0.1)) {
  names(x)<- c("BMI","age","gender","Race_Hispanic_origin",
               "Family_income_to_poverty_ratio","depression_score","alcohol")
  EN.fram = glmnet(x,y, alpha=0.5, family="binomial",
                   lambda=lambda_grid, data=pri_Q)
  cv.EN <- cv.glmnet(x,y, alpha=0.5, family="binomial", data=pri_Q)
  lambda_min_EN <- cv.EN$lambda.min
  lambda_1se_EN <- cv.EN$lambda.1se
  
}
cv.EN
'''
```



### Determine using categorical or continuous dp

$$
\log\left(\frac{p_{sleep~disorder}}{1-p_{sleep~disorder}}\right) = \beta_{k,0} + \beta_{k,1} depression_{i}
$$

### potential confounders

BMI, alcohol, gender, Race_Hispanic_origin, age, and Family_income_to_poverty_ratio appear to satisfy the causal definition of a confounder
```{r}
mod2 <- glm(sleep_trouble ~ BMI+alcohol+gender+Race_Hispanic_origin+age+Family_income_to_poverty_ratio+depression_score+avg_MAP,data = pri_Q, family = "binomial")
summary(mod2)
```

### potential effect modifier

```{r}
mod3 <- update(mod2,. ~ .+depression_score*BMI)
coef(summary(mod3))["BMI:depression_score", "Pr(>|z|)"]
mod3 <- update(mod2,. ~ .+depression_score*alcohol)
coef(summary(mod3))["alcohol:depression_score", "Pr(>|z|)"]
mod3 <- update(mod2,. ~ .+depression_score*gender)
coef(summary(mod3))["gender2:depression_score", "Pr(>|z|)"]
mod3 <- update(mod2,. ~ .+depression_score*Race_Hispanic_origin)
summary(mod3)
mod3 <- update(mod2,. ~ .+depression_score*age)
coef(summary(mod3))["age:depression_score", "Pr(>|z|)"]
mod3 <- update(mod2,. ~ .+depression_score*Family_income_to_poverty_ratio)
summary(mod3)
mod3 <- update(mod2,. ~ .+depression_score*avg_MAP)
coef(summary(mod3))["depression_score:avg_MAP", "Pr(>|z|)"]
mod3 <- update(mod2,. ~ .+depression_score*total_cholesterol)
coef(summary(mod3))["depression_score:total_cholesterol", "Pr(>|z|)"]
```

### Assess possible nonlinear effect of BMI, alcohol, Family_income_to_poverty_ratio, and age

```{r}
mod3 <- update(mod2,. ~ .+ I(BMI^2))
coef(summary(mod3))["I(BMI^2)", "Pr(>|z|)"]
mod3 <- update(mod2,. ~ .+ I(alcohol^2))
coef(summary(mod3))["I(alcohol^2)", "Pr(>|z|)"]
mod3 <- update(mod2,. ~ .+ I(age^2))
coef(summary(mod3))["I(age^2)", "Pr(>|z|)"]
mod4 <- update(mod3,. ~ .+ I(alcohol^2))
coef(summary(mod4))["I(alcohol^2)", "Pr(>|z|)"]
mod4 <- update(mod3,. ~ .+ I(Family_income_to_poverty_ratio^2))
coef(summary(mod4))["I(Family_income_to_poverty_ratio^2)", "Pr(>|z|)"]
mod4 <- update(mod3,. ~ .+I(avg_MAP^2))
coef(summary(mod4))["I(avg_MAP^2)", "Pr(>|z|)"]
mod4 <- update(mod3,. ~ .+I(total_cholesterol^2))
coef(summary(mod4))["I(total_cholesterol^2)", "Pr(>|z|)"]
```

$$
\log\left(\frac{p_{sleep~disorder}}{1-p_{sleep~disorder}}\right) = \beta_{k,0} + \beta_{k,1} depression_{i} + \beta_{k,2} BMI _{i} + \beta_{k,3} alcohol_{i} + \beta_{k,4} Race_{i} + \beta_{k,5} age_{i} + \beta_{k,6} age^2_{i}+ \beta_{k,7} income_{i} + \beta_{k,8} \text{avg_MAP}_{i} + \beta_{k,9} \text{total_cholesterol}_{i}
$$



```{r}
par(mfrow=c(1,1))
influencePlot(mod3,col="red")
influenceIndexPlot(mod3)
```

### GOF

Hosmer-Lemeshow test because we have a larger number of covariate patterns.

```{r}
library(ResourceSelection)
options(digits=7)
# Hosmer-Lemeshow Test
hoslem.test(mod3$y,fitted(mod3),g=10)
```

### ROC

```{r}
library(pROC)
predprob <- predict(mod3,type=c("response"))
roccurve <- roc(sleep_trouble ~ predprob, data = pri_Q)
plot(roccurve,col="red")
auc(roccurve)
```


The odds of ever having sleep trouble is estimated to be exp(0.1377799) = 1.147723 times the odds of ever having sleep trouble among poeple with a one-unit higher depression score on average, holding other covariates constant, according to these sample data. 


```{r}

```



